---
title: "J x K Design Allowing Simple Effects Inferences"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{J-K-Design_inference-on-simple-effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A vignette to do x y z

This is appropriate because...

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The packages

To use PsyR to calculate confidence intervals over contrast effects, we will utilise a workflow with functions from [afex](https://www.rdocumentation.org/packages/afex/versions/1.4-1) to implement ANOVA models, and [emmeans](https://rvlenth.github.io/emmeans/) to obtain estimated marginal means and the required contrast estimates. So let's load those packages, and any others that are helpful for our workflow.


```{r, loadpackages, message = FALSE, warning = FALSE}
library(afex)
library(emmeans)
library(PsyR)
library(tidyverse)
```

## The data

The `experience` dataframe contains the results of a between-subjects experiment examining the impact of years of experience (10, 3, less than 1) held by a therapist on therapy outcomes across therapy types (Behavioural, Cognitive, Psychotherapy and Group Therapy). A higher score on the dependent variable (outcome of therapy) is indicative of a better outcome. 

Let's load the data, then call the `head()` function to print the first 5 rows.

```{r loaddata}
data(experience)
head(experience, 5)
```

To learn more about the `experience` dataframe call the help function using `?df` to open its documentation.

```{r, whatexperience, message = FALSE, warning = FALSE, eval = FALSE}
?experience
```

Also, let's tidy up the data, labeling the levels of our independent variables for greater clarity.

```{r sortexperience}
experience <- experience %>%
  mutate(
    Experience = factor(
      Experience, 
      levels = c("10yrs", "3yrs", "<1yr")
    ),
    Therapy = factor(
      Therapy,
      levels = c("Behavioural", "Cognitive", "Psychotherapy", "Group")
    )
  )
```

Next, let's visualise our data. Doing so allows us to more intuitively
understand its layout.

```{r boxplotit, fig.width = 7, fig.height = 4, dev = 'svg'}
experience %>% 
  ggplot(aes(x = Therapy, y = Outcome, fill = Experience)) +
  geom_boxplot() + 
  theme_classic(base_size = 14)

```

##The Analysis

Let's prepare to run some statistical tests!

Here we will create the building blocks for first an ANOVA output, and
later our SCIs. To start, let's run our ANOVA and save it in the `mod`
object. Of note, defining mod in this way enables emmeans to compute
estimated marginal means from which to apply contrasts of interest.

```{r, aov}
# perform the statistical model
mod <- aov_ez(
  id = "SubjID",
  dv = "Outcome",
  data = experience,
  between = c("Experience", "Therapy")
)

emms <- emmeans(mod, c("Experience", "Therapy"))
emms
```

First, define the factor A (Experience of therapist) contrasts

```{r btwn_contrasts}
experience_contrasts = list(
  "lots - little" = c(0, 1, -1),
  "lots - some" = c(1, -1, 0)
)

therapy_contrasts = list(
  "evidence - woowoo" = c(0.5, 0.5, -0.5, -0.5),
  "behav - cog" = c(1, -1, 0, 0),
  "woo1 - woo2" = c(0, 0, 1, -1)
)  
```

## First, let's sort the A and A(b) and AB family

Calculate the main effects for factor A

```{r, factor_a_fx}
emms_exp <- emmeans(mod, c("Experience"))
exp_fx <- contrast(emms_exp, experience_contrasts, adjust = "bonferroni")
exp_fx
```

Now let's get the A(b) effects

```{r, A_b}

Exp_g_therapy <- emmeans(mod, ~ Experience | Therapy)
Ab_fx <- contrast(
  Exp_g_therapy, 
  method = experience_contrasts, 
  adjust = "bonferroni"
)

```

And last, the AB effects

```{r, AB}
interaction_fx <- contrast(
  emms,
  interaction = list(experience_contrasts, therapy_contrasts),
  adjust = "bonferroni"
)
interaction_fx
```

For a priori, bonferroni method:

Now we have our 3 sets of contrasts (all appropriately scaled).
The Bonferroni adjustment must be made at 
alpha = 0.5x2 for the A(B) family (or alpha=0.05x3, if using the all factorial model). 
alpha is divided by k, which = number of planned contrasts in the family.
Note that the A(B) family includes A simple effects contrasts, A main effects contrasts, and AB interaction contrasts.
v2 = N-JK (df error)
we can feed them into psyci to get our preferred confidence intervals.

ME (A) exp: k = 2 alpha = .1
A(b) exp @ each level of therapy k = 8, alpha = .1
AB interaction k = 6, alpha = .1
Thus k = 16, alpha = .1, for all families (as they are actually one family, as all non-independent)


```{r, get_cis}
all_fx <- c(exp_fx, Ab_fx, interaction_fx)
family_list = list("b") # to be documented
alphas = rep(list(2*.05), times=3)

psyci(
  model = mod, 
  contrast_tables = all_fx, 
  method = "bf", 
  family_list = list("b"),
  between_factors = list("Therapy", "Experience"), 
  alpha = alphas,
  independent = FALSE
)
```


Now do the post-hoc analysis including A(B) contrasts.

This is only Scheffe, as all are between.
Alphas = 1-(1-alpha)^m (where m = number of standard factorial families contained within the non-standard simple effects family).

To calculate nu1, where A has J levels and B has K levels

A(B) = K*(J-1) so A relates to Experience
B(A) = J*(K-1)



```{r get_cis_ph}
alphas = rep(list(1 - (1 - 0.05)^2), times = 3)
nu1 = length(unique(levels(experience$Therapy))) * 
      (length(unique(levels(experience$Experience))) - 1)

psyci(
  model = mod, 
  contrast_tables = all_fx, 
  method = "ph", 
  family_list = list("b"),
  between_factors = list("Therapy", "Experience"), 
  alpha = alphas,
  independent = FALSE, 
  nu1 = nu1
)
```

Now do a standard analysis with independent families, using the planned bonferonni procedure.

Calculate the main effects for factor B

```{r, factor_a_fx}
emms_exp <- emmeans(mod, c("Therapy"))
thrpy_fx <- contrast(emms_exp, therapy_contrasts)
thrpy_fx

stndrd_fx = list(exp_fx, thrpy_fx, interaction_fx)
```

Now get the CIs for the main effects and interactions, with no simple effects

```{r get_cis_no_sfx}
psyci(
  model = mod, 
  contrast_tables = stndrd_fx, 
  method = "bf", 
  family_list = list("b"), 
  between_factors = list("Therapy", "Experience"), 
  alpha = 0.05
)
```
